\Rpackage{MSnbase} input/output capabilities

### 2013-04-29 inst/doc/MSnbase-io.Rnw
add forword to io
export new readMSnSet2 function and new io section

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes \Rpackage{MSnbase}'s input and output 
  capabilites.
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, IO.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\input{Foreword.tex}

@@

\Rpackage{MSnbase}'s aims are to facilitate the reproducible analysis of 
mass spectrometry data within the \R environment, from raw data import and processing, 
feature quantification, quantification and statistical analysis of the results \citep{Gatto2012}.
Data import functions for several formats are provided and intermediate or final 
results can also be saved or exported. 
These capabilites are presented below.

@@

\paragraph{Quantitation data} Third party software can be used to generate 
quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
This data as well as any additional metadata can be imported with the 
\Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.

@@

\paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the metadata in the original 
\R object can not be included in the file. See \Rfunction{?writeMgfData} for details.

\paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
the \Rfunction{write.exprs} method. Feature metadata can be appended to the feature 
intenstiy values. See \Rfunction{?writeMgfData} for details.

@@

\section{Creating Robject{MSnSet} from text spread sheets}

@@

\subsection{A complete workflow}

The following section describes a workflow that uses three input files to 
create the \Robject{MSnSet}. These files respectively describe the quantitative 
expression data, the sample meta-data and the feature meta-data. 
It is taken from the \Rpackage{pRolco} tutorial and uses example files from 
the \Rpackage{pRolocdata} package.

@@

<<readCsvData0>>=
## The original data for replicate 1, available
## from the pRolocdata package
f0 <- dir(system.file("extdata", package = "pRolocdata"), 
          full.names = TRUE, pattern = "pr800866n_si_004-rep1.csv")
csv <- read.csv(f0)
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \section{Session information}
%% \label{sec:sessionInfo}

%% <<label=sessioninfo,results='asis',echo=FALSE,cache=FALSE>>=
%% toLatex(sessionInfo())
%% @

>>>>

%% Abstract and keywords %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vskip 0.3in minus 0.1in
\hrule
\begin{abstract}
  This vignette describes \Rpackage{MSnbase}'s input and output 
  capabilities.
\end{abstract}
\textit{Keywords}: Mass Spectrometry (MS), proteomics, infrastructure, IO.
\vskip 0.1in minus 0.05in
\hrule
\vskip 0.2in minus 0.1in
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{Foreword.tex}

@@

\Rpackage{MSnbase}'s aims are to facilitate the reproducible analysis of 
mass spectrometry data within the \R environment, from raw data import and processing, 
feature quantification, quantification and statistical analysis of the results \citep{Gatto2012}.
Data import functions for several formats are provided and intermediate or final 
results can also be saved or exported. 
These capabilities are presented below.

@@

\paragraph{Quantitation data} Third party software can be used to generate 
quantitative data and exported as a spreadsheet (generally comma or tab separated format). 
This data as well as any additional meta-data can be imported with the 
\Rfunction{readMSnSet} function. See \Rfunction{?readMSnSet} for details.

@@

\paragraph{Peak lists} \Robject{MSnExp} instances as well as individual spectra can be written 
as \texttt{mgf} files with the \Rfunction{writeMgfData} method. Note that the meta-data in the original 
\R object can not be included in the file. See \Rfunction{?writeMgfData} for details.

\paragraph{Quantitation data} Quantitation data can be exported to spreadsheet files with 
the \Rfunction{write.exprs} method. Feature meta-data can be appended to the feature 
intensity values. See \Rfunction{?writeMgfData} for details.

@@

\section{Creating \Robject{MSnSet} from text spread sheets}

@@

\subsection{A complete work flow}

The following section describes a work flow that uses three input files to 
create the \Robject{MSnSet}. These files respectively describe the quantitative 
expression data, the sample meta-data and the feature meta-data. 
It is taken from the \Rpackage{pRoloc} tutorial and uses example files from 
the \Rpackage{pRolocdata} package.

@@

<<readCsvData0, tidy = FALSE>>=
## The original data for replicate 1, available
## from the pRolocdata package
f0 <- dir(system.file("extdata", package = "pRolocdata"), 
          full.names = TRUE, 
          pattern = "pr800866n_si_004-rep1.csv")
csv <- read.csv(f0)
@

@@

\subsection{A shorter work flow}

The \Rfunction{readMSnSet2} function provides a simplified import workforce. 
It takes a single spreadsheet as input (default is \texttt{csv}) and extract the 
columns identified by \Robject{ecol} to create the expression data, while the 
others are used as feature meta-data. \Robject{ecol} can be a \Robject{character} with 
the respective column labels or a numeric with their indices. In the former case, 
it is important to make sure that the names match exactly. Special characters like 
\texttt{'-'} or \texttt{'('} will be transformed by \R into \texttt{'.'} when the 
\texttt{csv} file is read in.
Optionally, one can also specify a column to be used as feature names. 
Note that these must be unique to guarantee the final object validity.

<<readMSnSet2>>=
ecol <- paste("area", 114:117, sep = ".")
fname <- "Protein.ID"
eset <- readMSnSet2(f0, ecol, fname)
eset
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% section
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Session information}
\label{sec:sessionInfo}

<<label=sessioninfo,results='asis',echo=FALSE,cache=FALSE>>=
toLatex(sessionInfo())
@
